Data Prep

### 2024-03-28 scripts/00.0_dataPrep.qmd
Beginning to simplify with the new vultureUtils functions

Inputs:

-   `MB.LoginObject`

Outputs:

-   `data/dataPrep/inpa.feather`

```{r download-inpa}
inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
inpa <- methods::as(inpa, "data.frame")
inpa <- inpa %>%
  mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
         year = as.numeric(lubridate::year(timestamp))) %>%
  filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-09-15")) # cut this off at the same point as the ornitela data
write_feather(inpa, here::here("data", "dataPrep", "inpa.feather"))

length(unique(inpa$trackId))

# tidy up
rm(inpa)
gc()
```

@@

Inputs:

-   `MB.LoginObject`

Outputs:

-   `data/dataPrep/ornitela.feather`

```{r download-ornitela}
# Ornitela: download data from movebank (just a subset of the times for now)
minDate <- "2020-09-01 00:00"
maxDate <- "2023-09-15 11:59"
ornitela <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
write_feather(ornitela, here::here("data", "dataPrep", "ornitela.feather"))

length(unique(ornitela$trackId))

# tidy up
rm(ornitela)
rm(minDate)
rm(maxDate)
gc()
```

@@

*Inputs:*

-   `data/dataPrep/inpa.feather`

-   `data/dataPrep/ornitela.feather`

*Outputs:*

-   `data/dataPrep/joined.feather`

@@

# Will there be a problem joining these?
names(ornitela)[!(names(ornitela) %in% names(inpa))] # all names in "ornitela" are found in inpa
names(inpa)[!(names(inpa) %in% names(ornitela))] # many are not found in "ornitela" that are found in inpa

@@

rm(inpa)
rm(inpa_tojoin)
rm(ornitela)

## fix trackId
joined0 <- joined0 %>%
  mutate(trackId = as.character(trackId),
         trackId = case_when(trackId == "E03" ~ "E03w",
                             TRUE ~ trackId))

# number of unique individuals
length(unique(joined0$trackId))

# Add the Nili_id
ww <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, Movebank_id) %>%
  distinct()

# Are there any other individuals who have two different associated trackId's?
ww %>% filter(Movebank_id %in% joined0$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's

all(joined0$trackId %in% ww$Movebank_id) # false.

# Okay, let's investigate the ones that aren't included
joined0 %>%
  filter(!(trackId %in% ww$Movebank_id)) %>%
  pull(trackId) %>% unique() # looks like this is going to be an issue of character matching. Let's transform everything to lowercase and remove all characters and spaces

joined0$trackId <- str_replace_all(tolower(joined0$trackId), "\\.", "")
ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")

sort(unique(ww$mb_id))
unique(joined0$trackId)[!(unique(joined0$trackId) %in% ww$mb_id)]
joined0$trackId[joined0$trackId == "y01t60w"] <- "t60white"
joined0$trackId[joined0$trackId == "e86white"] <- "e86"
unique(joined0$trackId)[!(unique(joined0$trackId) %in% ww$mb_id)]
# Investigate the remaining ones:
joined0 %>% filter(!(trackId %in% ww$mb_id)) %>% 
  dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>%
  group_by(trackId, local_identifier, ring_id, dataset) %>% 
  summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
# I guess for now let's just make their trackId's into Nili_id's

joined <- left_join(joined0, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
                  by = c("trackId" = "mb_id"))
nrow(joined) == nrow(joined0) # good, same number of rows. (check that the join worked)
length(unique(joined$Nili_id))
# For any that didn't get Nili_ids assigned, make it the trackId
joined %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expected.
joined <- joined %>%
  mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
                             TRUE ~ Nili_id))
length(unique(joined$Nili_id))
write_feather(joined, here::here("data", "dataPrep", "joined.feather"))

# tidy up
rm(joined0)
rm(joined)
gc()
```

@@

*Inputs:*

-   `data/raw/whoswho_vultures_20230920_new.xlsx`

-   `data/dataPrep/joined.feather`

*Outputs:*

-   `data/dataPrep/removed_periods.feather`

@@

```{r remove-periods}
## annotate the data with periods to remove
periods_to_remove <- periods_to_remove %>%
  dplyr::select(Nili_id,
                remove_start,
                remove_end,
                reason) %>%
  mutate(across(c(remove_start, remove_end), .fns = function(x){
    lubridate::ymd(x)})) %>%
  dplyr::filter(!is.na(remove_end)) %>%
  group_by(Nili_id, reason) %>%
  # sequence of daily dates for each corresponding start, end elements
  dplyr::mutate(dateOnly = map2(remove_start, remove_end, 
                                seq, by = "1 day")) %>%
  # unnest the list column
  unnest(cols = c(dateOnly)) %>%
  # remove any duplicate rows
  distinct() %>%
  dplyr::select(-c(remove_start, remove_end)) %>%
  rename("status" = reason)

# Join to the original data
removal_annotated <- joined %>%
  left_join(periods_to_remove, by = c("Nili_id", "dateOnly")) %>%
  mutate(status = replace_na(status, "valid"))
nrow(removal_annotated) == nrow(joined) #T: good, same number of rows.

# Actually REMOVE the periods to remove...
removed_periods <- removal_annotated %>%
  filter(status == "valid")
nrow(removed_periods)
length(unique(removal_annotated$Nili_id))
nrow(joined) == nrow(removal_annotated) # true
nrow(removed_periods) != nrow(removal_annotated) # true

write_feather(removed_periods, here::here("data", "dataPrep", "removed_periods.feather"))

# tidy up
rm(periods_to_remove)
rm(joined)
rm(removal_annotated)
rm(removed_periods)
gc()
```

### Capture cages

This is Marta's code for identifying periods when the vultures were in the capture cages and removing days of data around those occasions.

To identify the capture dates, first we need to classify the roosts (now using the `get_roosts()` function). Then, if the bird roosted within 500m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset. This has been validated. This, however, does not work for the Carmel, because the roosts are within 500m of the cage and very often the birds roost on top of the cage without actually being inside it. So for the Carmel captures, we will use another protocol: if the birds were sleeping within 50m of the cage and it was a capture day (or 3 days before the release day), we consider the birds were captured and we remove the data.

Note: I'm calling the roosts data frame generated here `roosts0.Rda` because I want to distinguish it from the more final roost location *list* that we will generate further along in the cleaning process.

*Inputs:*

-   `data/dataPrep/removed_periods.feather`

-   `data/raw/capture_sites.csv`

-   `data/raw/all_captures_carmel_2010-2021.csv`

*Outputs:*

-   `data/dataPrep/roosts0.Rda`

-   `data/dataPrep/removed_captures.feather`

```{r}
removed_periods <- read_feather(here::here("data", "dataPrep", "removed_periods.feather"))
```

```{r remove-capture-cage-getroosts}
#| eval: false
# Identify and remove capture dates using Marta's code --------------------
roosts0 <- get_roosts_df(df = removed_periods, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
save(roosts0, file = here::here("data", "dataPrep", "roosts0.Rda"))

# tidy up
rm(roosts0)
```

```{r}
# Load information about the capture sites
capture_sites <- read.csv(here::here("data", "raw", "capture_sites.csv"))
load(here::here("data", "dataPrep", "roosts0.Rda"))
AllCarmelDates <- read.csv(here::here("data", "raw", "all_captures_carmel_2010-2021.csv"))
```

```{r remove-capture-cage}
# Identify the period of time during which the capture sites are open (when we need to do this exclusion)
start.day <- 01
start.month <- 08
end.day <-  30
end.month <- 11
distance <- 500 # distance, in meters, to calculate from the cage

# Subset the roost dataset with the start and end dates for the capture period
sub.roosts <- roosts0 %>%
  mutate(start_date = as.Date(paste(start.day, start.month, lubridate::year(date), sep="-"),
                              format="%d-%m-%Y"),
         end_date = as.Date(paste(end.day, end.month, lubridate::year(date), sep="-"),
                            format="%d-%m-%Y")) %>%
  filter(date >= start_date & date <= end_date)

unique(lubridate::month(sub.roosts$date)) # now only includes the fall months, which is the capture season.

# then we need to calculate the roost distance to each of the capture cages. if it is less than 500m, keep that line
crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat),
               nrow = nrow(sub.roosts), ncol = 2) # get roost locations as simple lat/long coordinates

DistanceMat <- matrix(ncol = nrow(capture_sites), nrow = nrow(crds))
colnames(DistanceMat) <- unique(capture_sites$name)

for(i in 1:nrow(crds)){ # for each roost point...
  DistanceMat[i,] <- round(geosphere::distm(crds[i,], capture_sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
  #print(i)
}

ClosestCaptureSite <- colnames(DistanceMat)[apply(DistanceMat,1,which.min)] # ID of closest capture site
ClosestCaptureDist <- apply(DistanceMat,1,min) # Distance from closest capture site

sub.roosts <- cbind(sub.roosts, ClosestCaptureSite, ClosestCaptureDist)
sub.roosts$Captured <- ifelse(sub.roosts$ClosestCaptureDist <= distance, "Yes", "No")

sub.captured <- subset(sub.roosts, Captured == "Yes") # get list of dates which the bird was inside the cage

sub.captured.dates <- subset(sub.captured,
                             select = c("Nili_id",
                                        "date",
                                        "ClosestCaptureSite",
                                        "ClosestCaptureDist",
                                        "Captured"))

## For Carmel--different protocol.
sub.captured.no.carmel <- subset(sub.captured.dates, ClosestCaptureSite != "Carmel")
sub.captured.carmel <- subset(sub.captured.dates, ClosestCaptureSite == "Carmel")

AllCarmelDates$Date <- as.Date(AllCarmelDates$Date, format = "%d/%m/%Y")

AllCarmelDates.1 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-1)))
AllCarmelDates.2 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-2)))
AllCarmelDates.3 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-3)))

AllCarmelDates.all <- rbind(AllCarmelDates, AllCarmelDates.1, AllCarmelDates.2, AllCarmelDates.3)

sub.captured.carmel <- sub.captured.carmel %>%
  mutate(known_capture = ifelse(date %in% AllCarmelDates.all$Date, 1, 0),
         Captured = ifelse(known_capture == 1 & ClosestCaptureDist <= 50, "yes", "no")) %>%
  filter(Captured == "yes") %>%
  dplyr::select(-c(known_capture))

names(sub.captured.no.carmel)
names(sub.captured.carmel)

sub.captured.dates <- rbind(sub.captured.no.carmel, sub.captured.carmel)

# We also need to exclude the day after the bird was captured
sub.captured.dates.1 <- sub.captured.dates
sub.captured.dates.1$date <- sub.captured.dates.1$date+1

sub.captured.dates <- rbind(sub.captured.dates, sub.captured.dates.1)
sub.captured.dates <- sub.captured.dates %>%
  dplyr::distinct(Nili_id, date, .keep_all = T)

# It all looks ok, so we can subset the dataset to exclude the capture dates
removed_captures <- removed_periods %>%
  left_join(sub.captured.dates, by = c("Nili_id", "dateOnly" = "date"))
nrow(removed_periods) == nrow(removed_captures) # should be TRUE. NOW we can filter.
removed_captures <- removed_captures %>%
  dplyr::filter(Captured != "Yes"|is.na(Captured)) # remove the individual*days when they were captured

# How many rows and individuals did we remove?
before <- nrow(removed_periods)
after <- nrow(removed_captures)
(propChange <- (before-after)/before) # approx. 1% of data removed.
length(unique(removed_periods$Nili_id)) == length(unique(removed_captures$Nili_id)) # no change in number of individuals! As expected.

removed_captures <- removed_captures %>%
  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))

write_feather(removed_captures, here::here("data", "dataPrep", "removed_captures.feather"))

# tidy up
rm(AllCarmelDates)
rm(AllCarmelDates.1)
rm(AllCarmelDates.2)
rm(AllCarmelDates.3)
rm(AllCarmelDates.all)
rm(capture_sites)
rm(crds)
rm(DistanceMat)
rm(removed)
rm(removed_captures)
rm(roosts0)
rm(sub.captured)
rm(sub.captured.carmel)
rm(sub.captured.dates)
rm(sub.captured.dates.1)
rm(sub.captured.no.carmel)
rm(sub.roosts)
rm(removed_periods)
rm(after)
rm(before)
rm(ClosestCaptureDist)
rm(ClosestCaptureSite)
rm(distance)
rm(start.day)
rm(start.month)
rm(end.day)
rm(end.month)
rm(propChange)

gc()
```

@@

*Inputs:*

-   `data/raw/whoswho_vultures_20230920_new.xlsx`

-   `data/dataPrep/removed_captures.feather`

*Outputs:*

-   `data/dataPrep/with_age_sex.feather`

```{r attach-age-sex}
#| message: false
removed_captures <- read_feather(here::here("data", "dataPrep", "removed_captures.feather"))
age_sex <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, birth_year, sex) %>%
  distinct()
```

```{r}
#| eval: false
with_age_sex <- removed_captures %>%
  dplyr::select(-c("sex")) %>%
  left_join(age_sex, by = "Nili_id")
nrow(with_age_sex) == nrow(removed_captures) # should be TRUE
length(unique(with_age_sex$Nili_id)) #201

write_feather(with_age_sex, here::here("data", "dataPrep", "with_age_sex.feather"))

# tidy up
rm(removed_captures)
rm(with_age_sex)
rm(age_sex)
gc()
```

## Clean data

@@

*Inputs:*

-   `data/raw/CutOffRegion.kml`

-   `data/dataPrep/with_age_sex.feather`

*Outputs:*

-   `data/dataPrep/cleaned.feather`

```{r}
mask <- sf::st_read(here::here("data", "raw", "CutOffRegion.kml"))
with_age_sex <- read_feather(here::here("data", "dataPrep", "with_age_sex.feather"))
```

```{r clean-data}
#| eval: false
# Clean the data
## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
cleaned <- vultureUtils::cleanData(dataset = with_age_sex, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) 
length(unique(cleaned$Nili_id)) # 172 individuals
# Fix time zone so dates make sense ---------------------------------------
## Overwrite the dateOnly column from the new times
cleaned <- cleaned %>%
  mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
         dateOnly = lubridate::date(timestampIsrael),
         month = lubridate::month(dateOnly),
         year = lubridate::year(dateOnly)) # XXX consider removing the downsample argument. Also this function is insanely slow and needs to be fixed, good grief.

# This produces a huge file because it's an SF file. We don't really need it to be sf at this stage. Let's remove that.
cleaned <- sf::st_drop_geometry(cleaned)

save(cleaned, file = here::here("data", "dataPrep", "cleaned.Rda"))

# tidy up
rm(mask)
rm(with_age_sex)
rm(cleaned)
gc()
```

@@

*Inputs:*

-   `data/dataPrep/cleaned.Rda`

*Outputs:*

-   `data/dataPrep/seasons_list.Rda`

-   `data/dataPrep/season_names.Rda`

```{r}
load(here::here("data", "dataPrep", "cleaned.Rda"))
```

```{r split-seasons}
#| eval: false
# Split into Marta's 3 seasons --------------------------------------------
cleaned <- cleaned %>%
  mutate(start_breeding = ifelse(month %in% c(1:5),
                                 paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
                                 paste(as.character(year), "-", "12", "-", "15", sep = "")),
         start_summer = ifelse(month == 12,
                               paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
                               paste(as.character(year), "-", "05", "-", "15", sep = "")),
         start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
         start_breeding = as.Date(start_breeding),
         start_summer = as.Date(start_summer),
         season = case_when(
           dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
           dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
           dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
  filter(dateOnly != "2023-09-15") # remove September 15 2023, because it turns out the season boundaries are supposed to be non-inclusive and this is the latest one.

with_seasons <- cleaned %>%
  mutate(age = year - birth_year, #XXX consider moving this up to the age/sex section. Why is it happening here?
         sex = case_when(sex == "f/m" ~ "u",
                         TRUE ~ sex)) %>%
  mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
  group_by(Nili_id, year, month) %>%
  mutate(age = ifelse(month == 2, min(age) + 1, age)) %>%
  mutate(age_group = case_when(age == 0 ~ "juv",
                               age >= 1 & age <= 4 ~ "sub",
                               age >= 5 ~ "adult",
                               TRUE ~ NA)) %>%
  mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
                                  season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
                                  season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
  # explicitly set the factor levels in temporal order
  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding", "2023_summer")),
         season = factor(season, levels = c("breeding", "summer", "fall")))

# remove columns that were used in the season calculation
with_seasons <- with_seasons %>%
  dplyr::select(-c(month, start_breeding, start_summer, start_fall))

@@

*Inputs:*

-   `data/dataPrep/seasons_list.Rda`
-   `data/dataPrep/season_names.Rda`

*Outputs:*

-   `data/dataPrep/removed_northern.Rda`

```{r}
load(here::here("data", "dataPrep", "seasons_list.Rda"))
load(here::here("data", "dataPrep", "season_names.Rda"))
```

@@

## Remove individuals not in the south
removed_northern <- map2(.x = seasons_list, 
                         .y = southern_indivs, ~.x 
                         %>% filter(Nili_id %in% .y))
save(removed_northern, file = here::here("data", "dataPrep", "removed_northern.Rda"))

@@

*Inputs:*

-   `data/dataPrep/removed_northern.Rda`
-   `data/dataPrep/season_names.Rda`

*Outputs:*

-   `data/dataPrep/low_fix_rate_indivs.Rda`

@@

*Inputs:*

-   `data/dataPrep/removed_northern.Rda`

-   `data/dataPrep/low_fix_rate_indivs.Rda`

@@

*Inputs:*

-   `data/dataPrep/downsampled_10min_forSocial.Rda`

*Outputs:*

-   `data/dataPrep/roosts.Rda`

@@

*Inputs:*

-   `data/dataPrep/removed_lfr.Rda`
-   `data/dataPrep/season_names.Rda`

*Outputs:*

-   `data/dataPrep/removed_nightime.Rda`

@@

*Inputs:*

-   `data/dataPrep/removed_nighttime.Rda`

*Outputs:*

-   `data/dataPrep/removed_lowppd.Rda`

@@

*Inputs:*

-   `data/dataPrep/removed_lowppd.Rda`
-   `data/dataPrep/season_names.Rda`

*Outputs:*

-   `data/dataPrep/removed_too_few_days.Rda`
-   `data/dataPrep/season_durations.Rda`

@@

*Inputs:*

-   `data/dataPrep/removed_too_few_days`

*Outputs:*

-   `data/dataPrep/elevs_z10.Rda`

-   `data/dataPrep/with_altitudes.Rda`

@@

*Inputs:*

-   `data/dataPrep/with_altitudes.Rda`

-   `data/dataPrep/season_names.Rda`

*Outputs:*

-   `data/dataPrep/downsampled_10min.Rda`

>>>>

```{r download-inpa}
inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
inpa <- methods::as(inpa, "data.frame")
inpa <- inpa %>%
  mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
         year = as.numeric(lubridate::year(timestamp))) %>%
  filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-09-15")) # cut this off at the same point as the ornitela data
write_feather(inpa, here::here("data", "dataPrep", "inpa.feather"))
```

@@

Data collected by the TAU-UCLA partnership under the NSF-BSF grant. Ornitela tags.

```{r download-ornitela}
# Ornitela: download data from movebank (just a subset of the times for now)
minDate <- "2020-09-01 00:00"
maxDate <- "2023-09-15 11:59"
ornitela <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
write_feather(ornitela, here::here("data", "dataPrep", "ornitela.feather"))
```

@@

@@

# Will there be a problem joining these?
length(names(ornitela)[!(names(ornitela) %in% names(inpa))]) # all names in "ornitela" are found in inpa
length(names(inpa)[!(names(inpa) %in% names(ornitela))]) # many are not found in "ornitela" that are found in inpa

@@

rm(inpa)
rm(inpa_tojoin)
rm(ornitela)
```

```{r join-nili-ids}
ww <- read_excel(here("data/raw/whoswho_vultures_20230920_new.xlsx"), 
                 sheet = "all gps tags")
ww_tojoin <- ww %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() # pull out just the names columns, nothing else, and remove any duplicates

# Prepare for join: are there any individuals in the `local_identifier` column of `joined0` that don't appear in the `Movebank_id` column of `ww_tojoin`?
problems <- joined0 %>% filter(!(local_identifier %in% ww_tojoin$Movebank_id)) %>% pull(local_identifier) %>% unique()
problems #let's check these against the who's who and see if we can make some reasonable changes.

## Fixes:
# Typo in the Movebank_id column of the who's who:
ww_tojoin <- ww_tojoin %>% mutate(Movebank_id = case_when(Movebank_id == "A65 Whiite" ~ "A65 White",
                                                          .default = Movebank_id))
# Fixes to joined0:
joined0 <- joined0 %>%
  mutate(local_identifier = case_when(local_identifier == "E86 White" ~ "E86",
                                      local_identifier == "E88 White" ~ "E88w",
                                      .default = local_identifier))

## Look for any remaining problems:
problems <- joined0 %>% filter(!(local_identifier %in% ww_tojoin$Movebank_id)) %>% pull(local_identifier) %>% unique()
problems #going to fix both of these afterward. E66 isn't listed in the Who's who at all, so we'll just call it "E66" in the Nili_id. The other one, Y01>T60 W, I've manually determined is Nili_id "tammy".

# join by movebank ID
joined <- left_join(joined0, ww_tojoin, 
                        by = c("local_identifier" = "Movebank_id"))
joined <- joined %>%
  mutate(Nili_id = case_when(is.na(Nili_id) & local_identifier == "E66 White" ~ "E66",
                             is.na(Nili_id) & local_identifier == "Y01>T60 W" ~ "tammy",
                             .default = Nili_id))

# Are there any remaining NA's for Nili_id?
nas <- joined %>% filter(is.na(Nili_id)) %>% pull(local_identifier) %>% unique()
length(nas) # yay, no more!

write_feather(joined, here("data/dataPrep/joined.feather"))
```

@@

@@

```{r remove-periods}
removed_periods <- vultureUtils::removeInvalidPeriods(dataset = joined, periodsToRemove = periods_to_remove)
save(removed_periods, file = here("data/dataPrep/removed_periods.Rda"))
```

## Clean the data based on filters for invalid data points (see `vultureUtils` package code for details)

```{r}
# warning: this takes a long time and uses a lot of memory!
data_cleaned <- vultureUtils::cleanData(dataset = removed_periods,
                     precise = F,
                     longCol = "location_long",
                     latCol = "location_lat",
                     idCol = "Nili_id",
                     report = F)
rm(joined)
gc()
save(data_cleaned, file = here("data/dataPrep/data_cleaned.Rda"))

(nrow(removed_periods)-nrow(data_cleaned))/nrow(removed_periods) # removed less than 1% of the points.
```

## Remove time spent in capture cages
```{r}
capture_sites <- read.csv(here("data/raw/capture_sites.csv"))
carmel <- read.csv(here("data/raw/all_captures_carmel_2010-2021.csv"))
removed_captures <- removeCaptures(data = data_cleaned, 
                                       captureSites = capture_sites, 
                                       AllCarmelDates = carmel, 
                                       distance = 500, idCol = "Nili_id")

(nrow(removed_captures)-nrow(data_cleaned))/nrow(data_cleaned) # did not remove any data. Should probably check on this...
save(removed_captures, file = here("data/dataPrep/removed_captures.Rda"))
```

@@

```{r attach-age-sex}
#| message: false
load(here("data/dataPrep/removed_captures.Rda"))
age_sex <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), 
                      sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, birth_year, sex) %>%
  distinct()

with_age_sex <- removed_captures %>%
  dplyr::select(-c("sex")) %>%
  left_join(age_sex, by = "Nili_id")
nrow(with_age_sex) == nrow(removed_captures) # should be TRUE
length(unique(with_age_sex$Nili_id)) #200

# tidy up
rm(removed_captures)
rm(age_sex)
gc()
```

## Mask data

@@

```{r region-masking}
mask <- sf::st_read(here("data/raw/CutOffRegion.kml"))

# XXX TEST
test <- sample_n(with_age_sex, 50000)
reMask <- T # XXX INCORPORATE THIS INTO THE FUNCTION!!
masked_test <- vultureUtils::inMaskFilter(dataset = test, mask = mask, inMaskThreshold = 0.33)
# XXX END TEST

# Region masking
data_masked <- vultureUtils::inMaskFilter(dataset = with_age_sex, mask = mask, inMaskThreshold = 0.33, crs = "WGS84") # this takes a long time!!
save(data_masked, file = here("data/dataPrep/data_masked.Rda")) # note that this saves a 3-element list. Need to get Ryan to fix this.#XXX
load(here("data/dataPrep/data_masked.Rda"))
length(data_masked$Nili_id)
data_masked_2 <- data_masked$dataset

# Fix time zone so dates make sense ---------------------------------------
## Overwrite the dateOnly column from the new times
data_masked_2 <- data_masked_2 %>%
  mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
         dateOnly = lubridate::date(timestampIsrael),
         month = lubridate::month(dateOnly),
         year = lubridate::year(dateOnly))
save(data_masked_2, file = here("data/dataPrep/data_masked_2.Rda")) # just the df #XXX fix

# tidy up
rm(mask)
rm(with_age_sex)
gc()
```

@@

```{r split-seasons}
load(here("data/dataPrep/data_masked_2.Rda")) # XXX fix

# Split into Marta's 3 seasons --------------------------------------------
split <- data_masked_2 %>%
  mutate(start_breeding = ifelse(month %in% c(1:5),
                                 paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
                                 paste(as.character(year), "-", "12", "-", "15", sep = "")),
         start_summer = ifelse(month == 12,
                               paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
                               paste(as.character(year), "-", "05", "-", "15", sep = "")),
         start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
         start_breeding = as.Date(start_breeding),
         start_summer = as.Date(start_summer),
         season = case_when(
           dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
           dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
           dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
  filter(dateOnly != "2023-09-15") # remove September 15 2023, because it turns out the season boundaries are supposed to be non-inclusive and this is the latest one.

with_seasons <- split %>%
  mutate(age = year - birth_year, #XXX consider moving this up to the age/sex section. Why is it happening here?
         sex = case_when(sex == "f/m" ~ "u",
                         TRUE ~ sex)) %>%
  mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
  group_by(Nili_id, year, month) %>%
  mutate(age = ifelse(month == 2, min(age) + 1, age)) %>%
  mutate(age_group = case_when(age == 0 ~ "juv",
                               age >= 1 & age <= 4 ~ "sub",
                               age >= 5 ~ "adult",
                               TRUE ~ NA)) %>%
  mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
                                  season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
                                  season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
  # explicitly set the factor levels in temporal order
  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding", "2023_summer")),
         season = factor(season, levels = c("breeding", "summer", "fall")))

@@

```{r}
load(here("data/dataPrep/seasons_list.Rda"))
load(here("data/dataPrep/season_names.Rda"))
```

@@

## Remove individuals not in the south
removed_northern <- map2(.x = seasons_list, 
                         .y = southern_indivs, ~.x 
                         %>% filter(Nili_id %in% .y))
save(removed_northern, file = here("data/dataPrep/removed_northern.Rda"))

@@

@@

@@

@@

@@

@@

@@

@@
